%*******************************************************
% Acknowledgements
%*******************************************************
\refstepcounter{dummy}
\pdfbookmark[0]{Acknowledgements}{Acknowledgements}
\chapter*{Acknowledgements}

At this point, I want to express my gratitude to everyone who has played an important role during my Ph.D. Their assistance, guidance, and support have been invaluable over the past five years. Without their help, I would not be where I am now.

First, I would like to thank my supervisor, Prof. Robert F. Wimmer-Schweingruber, for accepting me to join the group and work on the exciting projects - Chang'E-4/LND and Solar orbiter/EPD. He supported me in publishing these results in scientific journals and presenting them at multiple international conferences. He is the most knowledgeable scientist I have met. He is always patient whenever I bother him with basic and stupid questions, and he can clearly explain those complicated concepts. I enjoy discussing science with him. I hope one day I could made him proud of me. 

I am also grateful to Prof. Jingnan Guo, who recommended me to Bob five years ago while searching for a Ph.D. position. After I came to Kiel, she guided and helped me immensely on the LND projects. I hope to have opportunities to continue our collaborations in the future. Besides, I would like to thank Prof. Dressing Nina, who provided valuable instructions and comments on my first SEP paper of LND. I also learned a lot while working together on the other SEP projects. Most importantly, I appreciate the reference letter she wrote for the postdoc application to Caltech. That recommendation must be one of the critical factors that help me to be selected from multiple candidates.

Furthermore, thanks are given to my lovely colleagues working on LND and SOLO projects and in the Kiel group. Firstly, I would like to thank my officemate, Dr. Lars Berger, for his patience and kindness whenever I wanted to discuss. These discussions with him were one of the best moments that I have had in Kiel. Despite that, I would like to show special thanks to him for his encouragement while finishing my thesis in the last two months. Your encouragement made me survive. But I must tell the truth: I still can not enjoy your heavy metal music. Secondly, I would like to show my appreciation to Dr. Verena Heidrich-Meisner, Dr. Patrick K\"{u}hl, and Dr. Marquardt Johannes for their proofreading of this thesis. They provided valuable suggestions and vastly improved the thesis. I also want to thank the following colleagues for their help: the future Ph.D., Alexander Kollhoff, Chaoran Gu, Salman Khaksarighiri, Kr\"{o}hnke Henning, as well as Dr. Daniel Pacheco, Dr. Liu Yang, Dr. Johan von Forstner, Dr. Jia Yu, and Henning Lohf. 

I also want to thank my friends in Kiel, Jinru He, Lei Shao, Xuenan Li, and Xiuming Sun, who have always supported, cared for, and helped me during the tough COVID times. I will never forget the time we enjoyed the fried chicken wings from KFC on the special Tuesday and indulged in the Catan all night until the following day. Those tough but happy times have made me stronger and more confident.

I want to thank my parents - Aifang Wang and Fazhen Xu - for their understanding and support during my trip to complete the degree, which is a very long trip that I have never thought of before. You are the best parents in the world.

In particular, I want to express my deepest love and special gratefulness to my wife, Zhenlin Zhu. It has been 11 years since we first met. You are the most important person in my life, and I can't imagine the life without you. You were once my “light”, and I firmly believe you can always be my "light". Thanks for your support all the time. I am so lucky to have you.

Further acknowledgments to the online thesis writing and proofreading tools include Chatgpt, Grammarly, NotionAI, and DeepL. They are mighty in correcting grammar mistakes.
 

% At this point, I would like to thank everyone who supported me during the course of my Ph.D. studies and my work on MSL/RAD, Chang'E 4 LND, as well as Solar Orbiter EPD.

% First of all, I am grateful to my supervisor, Prof. Robert Wimmer-Schweingruber for the opportunity to work on these exciting projects, and his helpful advice. He also made it possible that these results could be published in scientific journals and presented at many international conferences. Sincere thanks also to Prof. Jingnan Guo, who supported my work since my bachelor's thesis, and who I wish all the best for her new position in China.

% Furthermore, thanks to all my colleagues in the three mission teams for their continued support and helpful discussions, including Zigong Xu, Alexander Kollhoff, and my officemate Christoph Terasa for the productive and enjoyable collaboration, e.g. on low- and high-level software for the Solar Orbiter mission, which will hopefully facilitate the EPD data analysis in the Kiel team for years to come, and to the rest of the Extraterrestrial Physics group at Kiel University. I also thank the group of Manuela Temmer and Astrid Veronig at the University of Graz and Mateja Dumbović at Hvar Observatory with whom I worked in close collaboration for many of the Forbush decrease studies, and who I enjoyed meeting regularly at the conferences in Vienna, Hvar and San Francisco.

% I additionally want to thank the bachelor and master students who supported my work during their Hiwi positions, Charlotte Büschel and Niklas Lundt, and three secondary school students, Joana Wanger, Lukas Abegg and Markus Arndt, who contributed to the data sets used in my studies during their internships in the ET group.

% I would like to thank Anne Fischer as well as Hanna Giese and Knud Schröter for their very thorough proofreading of this thesis and valuable suggestions, and my parents Kristina and Michael and my brother Julius for their moral support. 

% Last but not least, the data analysis presented in this thesis, the typesetting of the thesis itself and the generation of most of the figures were made possible by a number of open source software projects, including, but not limited to, the ones acknowledged below:

% \begin{refsection}[software.bib]
%   \nocite{*}
%   \newrefcontext[sorting=none]
%   \printbibliography[heading=none]
% \end{refsection}