Between September 4 and 10, 2017, there was a sudden increase in solar activity when Active Region 12673 produced four X-class flares, including the two strongest flares of Solar Cycle 24 (X9.3 on September 6 and X8.2 on September 10).
These solar events, associated with the release of \acp{SEP} as well as the eruption of multiple fast \acp{CME}, caused a \ac{GLE} of energetic particles on the surface of Earth (\ac{GLE} \# 72 in the \ac{GLE} database at \url{https://gle.oulu.fi/}) as well as on Mars, making it the first \ac{GLE} observed simultaneously on the surface of two different planets.
The \ac{SEP} event was very widespread (Earth and Mars had a longitudinal separation of $\sim$\SI{155}{\degree} at that time) and the largest \ac{GLE} seen on the surface of Mars with \ac{MSL}/\ac{RAD} so far. At Earth, the disturbances associated with the solar events caused, e.g., a suppression of critical high frequency (HF) radio communications systems \citep{Frissell-2018,Bland-2018} as well as navigation satellite systems such as GPS \citep{Berdermann-2018,Sato-2019}.
These major events have been studied in great detail, many of the articles concerning these events can be found in the special issues of \citet{SpaceWeather-2018-special-issue-September-event} as well as \citet{GRL-2018-special-issue-September-event}, where the latter is focusing on its impact on Mars.

While the September 6 flare had the stronger X-ray emission, the \acp{SEP} associated with the later September 10 flare is the one that caused the \ac{GLE} seen at Earth and Mars, due to the better magnetic connection. Three CMEs also erupted on September 9 and 10 from the same active region in similar directions ($\sim$\SI{115}{\degree} in \ac{HEE} coordinates), the last one with an extremely high speed of more than \SI{2600}{\kilo\meter\per\second}. This fast and wide CME quickly merged with the two previous ones and formed an intense interplanetary shock, propagated outward and was observed \textit{in situ} at both Earth and Mars on September 12 and 13, respectively. While the \ac{SEP} event at Mars was still in its declining phase, the shock arrival caused an intense \ac{FD} at Mars with an amplitude of \SI{15}{\percent}, the largest \ac{FD} seen by \ac{RAD} to date. This decrease below the pre-\ac{SEP} levels was sustained over 5 days and then gradually recovered over the course of several weeks.

The following two articles \citep{Zeitlin-2018,Guo-2018} study the effects of the September 10 events on Mars as measured by \ac{MSL}/\ac{RAD} and explain these observations by modeling the \ac{SEP} event and the three \acp{CME}. \citet{Zeitlin-2018} report the dosimetric quantities on the Martian surface, emphasizing that the increased dose and dose equivalent rates during the \ac{GLE} on Mars are almost canceled out by the long-lasting \ac{FD} following it. Thus, in the case of a long-stay mission scenario, the increased radiation exposure due to the September event would have been insignificant for astronauts on Mars despite the 2- to 3-fold increase during the peak of the \ac{SEP} event. However, as Mars was not particularly well connected to the active region, the impact of the \ac{SEP} event could have been much larger, so this conclusion should not be generalized to all major solar events at Mars. In \citet{Guo-2018}, the propagation of the 3 \acp{ICME} towards Earth and Mars is studied in more detail. The initial parameters of the \acp{CME} close to the Sun are reconstructed using \ac{GCS} fitting, and then the kinematics of the propagation are calculated using the empirical \ac{DBM} \citep{Vrsnak-2007,Vrsnak-2013}. This is one of the first instances where \ac{DBM} is used to predict the merging of multiple \acp{ICME} --- for this process, a simple conservation of momentum is assumed, and with a sensible choice of the drag parameter $\gamma$, it fits the observed arrival times at Earth and Mars. The results are also compared to more sophisticated \ac{MHD} simulations performed for the same event.

\input{chapters/pub05_zeitlin2018}
\input{chapters/pub06_guo2018}